### 2021-02-17 outputs/paper/data_simulation.Rmd
stitching together simulation work and tidying file names

---
title: "Data for the Ontario Restaurant Closures Experiment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
subtitle: "Simulation Methods and Parameterization Research"
abstract: "This document explains how data for the Ontario restaurant closures experiment was simulated, including the research that was used for parameterization and the methods used to randomize data."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in projects

```

```{r}
# read in the data
# NOTE: scripts 01_scrape_health_depts, 04_clean_restaurant_list_csvs, and 02_sampling_frame must have been run already

table_for_surveys <- read_csv(here::here("outputs/table_for_surveys.csv"))

survey_1 <- table_for_surveys %>%
  select(ID, group, verify) %>%
  arrange(ID)

# count the numbers for treat, control, and total
num_treat <- first(count(survey_1 %>%
                           filter(group == "treatment")))

num_control <- first(count(survey_1 %>% 
                             filter(group == "control")))

num_total <- first(count(survey_1))

survey_2 <- survey_1

survey_1
```

# Survey questions that require simulated data

# Type of service provided
```{r}
# Read in data
yelp_data <- read_csv(here("inputs/data/yelp_restaurants_ontario.csv")) %>% 
  na.omit() # Remove all NAs
total_restaurants <- count(yelp_data)

# Check proportion of RestaurantsTakeOut
takeout <- yelp_data %>% 
  select(RestaurantsTakeOut)
summary(takeout)
```
```{r}
takeout_percentage <- takeout %>%
  group_by(RestaurantsTakeOut) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

takeout_percentage
```
```{r}
# Check proportion of RestaurantsTableService
tableserv <- yelp_data %>% 
  select(RestaurantsTableService)
summary(tableserv)
```
```{r}
tableserv_percentage <- tableserv %>%
  group_by(RestaurantsTableService) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

tableserv_percentage
```

```{r}
# Calculate invalid responses (neither dine-in nor takeout)
invalid <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == FALSE) %>% 
  count()

valid_restaurants <- total_restaurants-invalid

# Calculate service type percentages
takeout_only <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

dinein_only <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == FALSE) %>% 
  count()/valid_restaurants

both <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

service_percentage<- data.frame(
  service = c("takeout", "dinein", "both"),
  percentage = c(takeout_only$n, dinein_only$n, both$n), options(digits = 2)
)

 service_percentage
```
According to Ontario restaurant data from Yelp (cite, KSondjaja source), approximately 6% of restaurants
offer dine in service, 32.5% offer takeout, and 61.5% offer both.

## Survey 1
```{r}
set.seed(19)
survey_1 <- survey_1 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))
```

## Survey 2
```{r}
set.seed(19)
survey_2 <- survey_2 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))
```

# Demographic traits of owners

https://www.ic.gc.ca/eic/site/061.nsf/eng/h_03115.html
0.5% are owned by persons with disabilities

## Survey 1
```{r}
set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(disability = sample(c("yes", "no"),
                             size = num_total,
                             prob = c(0.005, .995),
                             replace = TRUE))

survey_1 <- survey_1 %>%

```

## Survey 2

Based on:https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310028101
there seems to be no significant difference in business losses for disabled owners.

```{r}
set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(disability = sample(c("yes", "no"),
                             size = num_total,
                             prob = c(0.005, .995),
                             replace = TRUE))

survey_2 <- survey_2 %>%
```

# Shutdowns

# Survey 2

https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310027601
43.2% shut down temporarily

https://nypost.com/2020/06/30/more-than-half-of-us-restaurants-closed-due-to-covid-19-study/
3% closed permanently - but that was over far more than two weeks.

```{r}
set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                            size = num_treat, 
                                                            prob = c(0.558, 0.432, 0.01),
                                                            replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                          size = num_control, 
                                                          prob = c(.99, 0.005, 0.005),
                                                          replace = TRUE))
) %>%
  arrange(ID)

```

# Revenue

## Survey 1

From https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310022201&pickMembers%5B0%5D=2.1&pickMembers%5B1%5D=3.423
25,836 employers

From https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310022301&pickMembers%5B0%5D=2.423
6,968 non-employers

So 32,804 businesses in 2019

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2110017101&pickMembers%5B0%5D=1.7&pickMembers%5B1%5D=2.2&cubeTimeFrame.startYear=2015&cubeTimeFrame.endYear=2019&referencePeriods=20150101%2C20190101
full-service restaurants have 13,456,600,000 in revenue in 2019.
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2110017101&pickMembers%5B0%5D=1.7&pickMembers%5B1%5D=2.3&cubeTimeFrame.startYear=2015&cubeTimeFrame.endYear=2019&referencePeriods=20150101%2C20190101
limited-service eating places have 14,082,700,000 in revenue in 2019.

so averaging 839,510 in annual revenue.

69,959 average per month.

From http://www.mbel.io/2019/08/23/kaggle-restaurant-revenue-prediction/
we can see the distribution looks like an F distribution, or a log normal distribution.
rf(n, df1, df2, ncp) = rf(num_rest, 10, 5) should have a mean of 5/3
so we would multiply by 41,975 to get the mean where we want.

```{r}
set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(revenue = 41975 * rf(num_total,10,5))
```

## Survey 2

need to account for total and partial shutdowns: assume that total shutdowns cut revenue 75%, and partial cut revenue 50%

from https://www.eater.com/2020/3/24/21184301/restaurant-industry-data-impact-covid-19-coronavirus
revenues went from +5 to -35 when the shutdowns started. loss of -38. 
but takeout only places need -0
dine-in only places -100*14/31 = -45
and combo places need -38*14/31 = -17

https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310028101
all business had a median decrease of about 15%
minority owned businesses had a median decrease of about 25%.
Decrease of about 12%.

```{r}
set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(revenue = 41975 * rf(num_total,10,5)) %>%
  mutate(factor = abs(rnorm(num_total, 1, .05))) %>%
  mutate(revenue = factor * revenue) %>%
  select(-factor)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(shutdown == "none") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(shutdown == "temporary") %>%
      mutate(revenue = 0.50 * revenue)
    ,
    survey_2 %>%
      filter(shutdown == "permanent") %>%
      mutate(revenue = 0.25 * revenue)
  ) %>%
  arrange(ID)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(service_type == "dinein" & group == "treatment") %>%
      mutate(revenue = 0.55 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "both" & group == "treatment") %>%
      mutate(revenue = 0.83 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "takeout" & group == "treatment") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(group == "control")
  ) %>%
  arrange(ID)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(minority == "yes" & group == "treatment") %>%
      mutate(revenue = 0.88 * revenue)
    ,
    survey_2 %>%
      filter(minority == "no" & group == "treatment")
    ,
    survey_2 %>%
      filter(group == "control")
  )

```

# Number of employees
LALA
## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2
```{r}
survey_2 <- survey_2 %>%
```

# Applications for government assistance
LORENA
## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2
```{r}
survey_2 <- survey_2 %>%
```

\newpage

# References