### 2021-02-18 outputs/paper/data_simulation.Rmd
data simulation done, all in-text references in data_simulation, first test knits run on both paper and data_simulation

---
title: "Data for the Ontario Restaurant Closures Experiment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
subtitle: "Simulation Methods and Parameterization Research"
abstract: "This document explains how data for the Ontario restaurant closures experiment was simulated, including the research that was used for parameterization and the methods used to randomize data."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in projects

```

```{r surveyframe}
# read in the data
# NOTE: scripts 01_scrape_health_depts, 04_clean_restaurant_list_csvs, and 02_sampling_frame must have been run already

table_for_surveys <- read_csv(here::here("outputs/table_for_surveys.csv"))

survey_1 <- table_for_surveys %>%
  select(ID, group, verify) %>%
  arrange(ID)

# count the numbers for treat, control, and total
num_treat <- first(count(survey_1 %>%
                           filter(group == "treatment")))

num_control <- first(count(survey_1 %>% 
                             filter(group == "control")))

num_total <- first(count(survey_1))

survey_2 <- survey_1

survey_1
```

# Survey questions that require simulated data

# Type of service provided

```{r}
# Read in data
yelp_data <- read_csv(here("inputs/data/yelp_restaurants_ontario.csv")) %>% 
  na.omit() # Remove all NAs
total_restaurants <- count(yelp_data)

# Check proportion of RestaurantsTakeOut
takeout <- yelp_data %>% 
  select(RestaurantsTakeOut)
summary(takeout)
```
```{r}
takeout_percentage <- takeout %>%
  group_by(RestaurantsTakeOut) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

takeout_percentage
```
```{r}
# Check proportion of RestaurantsTableService
tableserv <- yelp_data %>% 
  select(RestaurantsTableService)
summary(tableserv)
```
```{r}
tableserv_percentage <- tableserv %>%
  group_by(RestaurantsTableService) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

tableserv_percentage
```

```{r}
# Calculate invalid responses (neither dine-in nor takeout)
invalid <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == FALSE) %>% 
  count()

valid_restaurants <- total_restaurants-invalid

# Calculate service type percentages
takeout_only <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

dinein_only <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == FALSE) %>% 
  count()/valid_restaurants

both <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

service_percentage<- data.frame(
  service = c("takeout", "dinein", "both"),
  percentage = c(takeout_only$n, dinein_only$n, both$n), options(digits = 2)
)

 service_percentage
```
According to Ontario restaurant data from Yelp [@yelp_dataset], approximately 6% of restaurants
offer dine in service, 32.5% offer takeout, and 61.5% offer both.

## Survey 1
```{r}
set.seed(19)
survey_1 <- survey_1 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))

survey_1
```

## Survey 2
```{r}
set.seed(19)
survey_2 <- survey_2 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))

survey_2
```

# Demographic traits of owners

0.5% of small to medium enterprises in Canada are owned by persons with disabilities [@ownershipdemos].

## Survey 1
```{r}
set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(disability = sample(c("yes", "no"),
                             size = num_total,
                             prob = c(0.005, .995),
                             replace = TRUE))

survey_1
```

## Survey 2
```{r}
set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(disability = sample(c("yes", "no"),
                             size = num_total,
                             prob = c(0.005, .995),
                             replace = TRUE))

survey_2
```

```{r}
### Load census dataset by Health Regions ###
### Download link: https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/download-telecharger/comp/GetFile.cfm?Lang=E&FILETYPE=CSV&GEONO=058 ###
census_2016 <- read.csv(here("inputs", "data", "census_2016.csv"))
census_2016 <- janitor::clean_names(census_2016)

# Narrowed down categories
demographic_info <- c(1, 8, 1324, 1290, 1917)

### Function to get data from specific regions and rows ###

get_region_data <- function(y){
  census_2016 %>%
    filter(geo_code_por == y,
           member_id_profile_of_health_regions_2247 %in% demographic_info) %>%
    select(dim_profile_of_health_regions_2247, member_id_profile_of_health_regions_2247, dim_sex_3_member_id_1_total_sex, dim_sex_3_member_id_2_male, dim_sex_3_member_id_3_female)
}

### Get data from the randomly selected Health Regions ###
# GEO_CODE indicates public health regions, taken from Census 2016 links:
#https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/search-recherche/lst/results-resultats.cfm?Lang=E&TABID=1&G=1&Geo1=&Code1=&Geo2=&Code2=&GEOCODE=35&type=0

ontario <- get_region_data(35)

haliburton <- get_region_data(3535)
hamilton <- get_region_data(3537)
algoma <- get_region_data(3526)
simcoe_muskoka <- get_region_data(3560)
timiskaming <- get_region_data(3563)
windsor_essex <- get_region_data(3568)

northwest <- get_region_data(3549)
# southwestern <- get_region_data(3502)
# Southwestern is a newer health unit. it was created by almagamating oxford and elgin-st. thomas units:
oxford <- get_region_data(3552)
elgin <- get_region_data(3531)
waterloo <- get_region_data(3565)
durham <- get_region_data(3530)
sudbury <- get_region_data(3561)
brant <- get_region_data(3527)

### Put together demographic info in one table ###

populationss <- c("Total_Population", "Indigenous_Population_25%_sample", "Visible_Minority_25%_sample", "Accommodation_and_Food_Services_25%_sample", "Total_Population_Women")

get_pop_info <- function(x){
  c(as.numeric(x$dim_sex_3_member_id_1_total_sex[x$dim_profile_of_health_regions_2247 == "Population, 2016"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex[x$dim_profile_of_health_regions_2247 == "Aboriginal identity"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex[x$dim_profile_of_health_regions_2247 == "Total visible minority population"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex[x$dim_profile_of_health_regions_2247 == "72 Accommodation and food services"]),
    as.numeric(x$dim_sex_3_member_id_3_female[x$dim_profile_of_health_regions_2247 == "Total - Age groups and average age of the population - 100% data"]))
}

ontario_pop <- get_pop_info(ontario)

haliburton_pop <- get_pop_info(haliburton)
algoma_pop <- get_pop_info(algoma)
hamilton_pop <- get_pop_info(hamilton)
windsor_essex_pop <- get_pop_info(windsor_essex)
simcoe_muskoka_pop <- get_pop_info(simcoe_muskoka)
timiskaming_pop <- get_pop_info(timiskaming)

brant_pop <- get_pop_info(brant)
sudbury_pop <- get_pop_info(sudbury)
#southwestern_pop <- get_pop_info(southwestern)
oxford_pop <- get_pop_info(oxford)
elgin_pop <- get_pop_info(elgin)
northwest_pop <- get_pop_info(northwest)
waterloo_pop <- get_pop_info(waterloo)
durham_pop <- get_pop_info(durham)

populations <- bind_cols(populationss, ontario_pop, haliburton_pop, algoma_pop, hamilton_pop, windsor_essex_pop, simcoe_muskoka_pop, timiskaming_pop, brant_pop, sudbury_pop, oxford_pop, elgin_pop, northwest_pop, waterloo_pop, durham_pop)

colnames(populations) <- c("Info", "Ontario", "Haliburton", "Algoma", "Hamilton", "Windsor_Essex", "Simcoe_Muskoka", "Timiskaming",
                           "Brant", "Sudbury", "Oxford", "Elgin", "Northwest", "Waterloo", "Durham")

populations <- populations %>%
  mutate(Southwestern = Oxford + Elgin) %>%
  mutate(total_treat = Haliburton + Algoma + Hamilton + Windsor_Essex + Simcoe_Muskoka + Timiskaming,
         total_control = Brant + Sudbury + Northwest + Waterloo + Durham + Southwestern) %>%
  select(-Oxford, -Elgin)

populations
```

```{r}
populations_split <- matrix(ncol=4, nrow=5)
populations_split[,1] <- c("Total_Population", "Indigenous_Population_25%_sample", "Visible_Minority_25%_sample", "Accommodation_and_Food_Services_25%_sample", "Total_Population_Women")
populations_split[,2] <- ontario_pop

total_pop_treatment <- sum(populations[1,c(3:8)])
indigenous_treatment <- sum(populations[2,c(3:8)])
minority_treatment <- sum(populations[3,c(3:8)])
food_services_treatment <- sum(populations[4,c(3:8)])
women_treatment <- sum(populations[5,c(3:8)])

treatment <- c(total_pop_treatment, indigenous_treatment, minority_treatment, food_services_treatment, women_treatment)
populations_split[,3] <- treatment

total_pop_control <- sum(populations[1,c(9:14)])
indigenous_control <- sum(populations[2,c(9:14)])
minority_control <- sum(populations[3,c(9:14)])
food_services_control <- sum(populations[4,c(9:14)])
women_control <- sum(populations[5,c(9:14)])

control <- c(total_pop_control, indigenous_control, minority_control, food_services_control, women_control)
populations_split[,4] <- control

populations_split <- as.data.frame(populations_split)
colnames(populations_split) <- c("Info", "Ontario", "Treatment", "Control")

populations_split
```

```{r}
### Get percentage proportions of demographic groups of interest ###

populations_split_percentage <- matrix(ncol=4, nrow=5)

for(i in 2:length(populations_split)){
  for(j in 1:5){
    populations_split_percentage[j,i] <- round(as.numeric(populations_split[j,i])/as.numeric(populations_split[1,i]), digits=3)
  }
}

#populations_percentage[1, 2:4] <- c(1, 1, 1)
populations_split_percentage[,1] <- c("Total_Population", "Indigenous_Population_25%_sample", "Visible_Minority_25%_sample", "Accommodation_and_Food_Services_25%_sample", "Total_Population_Women")

populations_split_percentage <- as.data.frame(populations_split_percentage)
colnames(populations_split_percentage) <- c("Info", "Ontario", "Treatment", "Control")

populations_split_percentage
```

We established expected percentages for the treatment and control regions based on census data [@census_2016].

15.6% of business owners are women in Canada [@ownershipdemos].

Visible minorities are 25% of the population [@visibleminority].

12.2% of business owners are visible minorities [@ownershipdemos].

Therefore, visible minorities are 0.49 as likely to own a business.
Therefore, we can estimate that 0.058 of restaurant owners are a visible minority in the treatment group, and 0.081 in the control group.

Indigenous people have 3.7% service business ownership as opposed to 15.3% in the reference population [@businessesper]. They are 0.24 times as likely to own a service business in Ontario.
Therefore, we can estimate that 0.009 of business owners in the treatment group are Indigenous, and 0.011 in the control group.

## Survey 1
```{r}
set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(woman = sample(c("yes","no","nonanswer"), size = num_total, prob = c(0.154, 0.836, 0.01), replace = TRUE))

set.seed(19893)
survey_1 <- bind_rows(
  survey_1 %>%
    filter(group == "treatment") %>%
    mutate(indigenous = sample(c("yes", "no", "nonanswer"), size = num_treat, prob = c(0.009, 0.981, 0.01), replace = TRUE))
  ,
  survey_1 %>%
    filter(group == "control") %>%
   mutate(indigenous = sample(c("yes", "no", "nonanswer"), size = num_control, prob = c(0.011, 0.979, 0.01), replace = TRUE))
)

set.seed(19893)
survey_1 <- bind_rows(
  survey_1 %>%
    filter(group == "treatment", indigenous != "yes") %>%
    mutate(visible = sample(c("yes", "no", "nonanswer"), 
                               size = first(count(survey_1 %>% filter(group == "treatment", indigenous != "yes"))), 
                               prob = c(0.057, 0.933, 0.01), replace = TRUE))
  ,
  survey_1 %>%
    filter(group == "control", indigenous != "yes") %>%
   mutate(visible = sample(c("yes", "no", "nonanswer"), 
                              size = first(count(survey_1 %>% filter(group == "control", indigenous != "yes"))), 
                              prob = c(0.08, 0.91, 0.01), replace = TRUE))
  ,
  survey_1 %>%
    filter(indigenous == "yes") %>%
    mutate(visible = "no")
)

survey_1
```

## Survey 2

```{r}
set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(woman = sample(c("yes","no","nonanswer"), size = num_total, prob = c(0.154, 0.836, 0.01), replace = TRUE))

set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(indigenous = sample(c("yes", "no", "nonanswer"), size = num_treat, prob = c(0.009, 0.981, 0.01), replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
   mutate(indigenous = sample(c("yes", "no", "nonanswer"), size = num_control, prob = c(0.011, 0.979, 0.01), replace = TRUE))
)

set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment", indigenous != "yes") %>%
    mutate(visible = sample(c("yes", "no", "nonanswer"), 
                               size = first(count(survey_2 %>% filter(group == "treatment", indigenous != "yes"))), 
                               prob = c(0.057, 0.933, 0.01), replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control", indigenous != "yes") %>%
   mutate(visible = sample(c("yes", "no", "nonanswer"), 
                              size = first(count(survey_2 %>% filter(group == "control", indigenous != "yes"))), 
                              prob = c(0.08, 0.91, 0.01), replace = TRUE))
  ,
  survey_1 %>%
    filter(indigenous == "yes") %>%
    mutate(visible = "no")
)

survey_2
```

# Shutdowns

# Survey 2

43.2% of restaurants shut down temporarily due to the pandemic [@covid19impactonbusiness].

3% closed of restaurants in the US closed permanently between March and June [@restaurantclosures]. This period was about six times as long as our two week closure.

```{r}
set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                            size = num_treat, 
                                                            prob = c(0.558, 0.432, 0.01),
                                                            replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                          size = num_control, 
                                                          prob = c(.99, 0.005, 0.005),
                                                          replace = TRUE))
) %>%
  arrange(ID)

survey_2
```

# Revenue

## Survey 1

In 2019, there were 25,836 restaurants and eating-places that were employers [@businesseswithemployees] and 6,968 restaurants that were not employers [@businesseswithoutemployees] in Canada, for a total of 32,804 restaurants.

In the same year, full-service restaurants had \$13,456,600,000 in revenue, and limited-service eating places had \$14,082,700,000 in revenue [@foodservicessummary].

From this, we can calculate an estimated \$69,959 average revenue per month.

From an analysis of Kaggle restaurant data [@kagglerestaurantdata], we can see that restaurant revenue looks like an F distribution.
The distribution rf(n, df1, df2, ncp) = rf(num_rest, 10, 5) gives an approximate shape. Using the f-distribution formula, we know the distribution has a mean of 10/9, so we would multiply by 62,963 to get the mean to 69,959.

```{r}
set.seed(19894)
survey_1 <- survey_1 %>%
  mutate(revenue = 62963 * rf(num_total,20,20))

survey_1
```

## Survey 2

First, the revenues from the first survey will have a randomized factor (normal distribution with a mean of 1 and a standard distribution of 0.1) to add some random variance between months.

Second, places that are closed permanently will be assumed to experience -75% average revenue, and temporary shutdowns will be assumed to experience -50% revenue.

Revenues went to -35% during the first lockdown wave in March [@bythenumbers]. We will assume that these numbers are accurate for a dine-in/takeout restaurant that is experiencing a takeout-only closure. They are closed 14 out of 31 days, so we will assume average losses of -17%.

We will assume that takeout-only restaurants will not have losses on average, and dine-in only places will have -100% revenue on average for the days they are closed. With closures 14 out of 31 days in the month, dine-in only will have average losses of -45%.

All businesses had a median decrease of -15%, but visible minority-owned businesses had a median decrease of -25% [@aug20192020]. Therefore, an additional -12% will be applied to visible minority-owned restaurants.

There seems to be no significant difference in business losses for disabled owners or female business owners [@aug20192020].

```{r}
set.seed(19894)
survey_2 <- survey_2 %>%
  mutate(revenue = 62963 * rf(num_total,20,20)) %>%
  mutate(factor = abs(rnorm(num_total, 1, .1))) %>%
  mutate(revenue = factor * revenue) %>%
  select(-factor)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(shutdown == "none") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(shutdown == "temporary") %>%
      mutate(revenue = 0.50 * revenue)
    ,
    survey_2 %>%
      filter(shutdown == "permanent") %>%
      mutate(revenue = 0.25 * revenue)
  ) %>%
  arrange(ID)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(service_type == "dinein" & group == "treatment") %>%
      mutate(revenue = 0.55 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "both" & group == "treatment") %>%
      mutate(revenue = 0.83 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "takeout" & group == "treatment") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(group == "control")
  ) %>%
  arrange(ID)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter((visible == "yes" | indigenous == "yes") & group == "treatment") %>%
      mutate(revenue = 0.88 * revenue)
    ,
    survey_2 %>%
      filter(visible == "no" & group == "treatment")
    ,
    survey_2 %>%
      filter(group == "control")
  )

survey_2
```

# Number of employees

$82,000 in yearly sales (gross revenue) per full-time equivalent employee [@restfact]

Therefore 6,833 monthly revenue per full-time equivalent employee.

In January 2021, there were 2187300 full time sales and service employees in Canada and 1164000 part time sales and service employees [@hourswages].
Therefore there are 1.9 full time employees per part time employee.
Part time employees in service and sales averaged 16.6 hours per week. We will assume that 2 PTE = 1 FTE, or 1 PTE = 0.5 FTE.
Therefore, for every 1.9 + 1 employees, there are 1.9 + 1*0.5 FTE. For every 2.9 employees, there are 2.4 FTE.

2.4 FTE -> 1.9 FT + 1 PT.
0.79 = 1.9 / 2.4 = 1.9 / (1.9 + 0.5)
0.42 = 1 / 2.4 = 0.42 = 1 / (1.9 + 0.5)

## Survey 1
```{r}
set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(fte = revenue / 6833) %>%
  mutate(ft = round(fte * 0.79)) %>%
  mutate(pt = round(fte * 0.42)) %>%
  select(-fte)

survey_1 <- bind_rows(
  survey_1 %>% filter(ft < 1) %>%
    mutate(ft = 1),
  survey_1 %>% filter(ft >= 1)
)

survey_1
```

## Survey 2

For the second survey, we will assume that treatment restaurants shifted towards part time random amounts-- that is, the number of full time employees per part time employee dropped from 1.9. X is some number less than or equal to 1.9.
x / (x + 0.5)
1 / (x + 0.5)

```{r}
set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(fte = revenue / 6833) %>%
  mutate(ptincrease = 1.9 * (1 - abs(rnorm(num_total, 0, sd = 0.1)))) %>%
  mutate(ft = round(fte * (ptincrease / (ptincrease + 0.5)))) %>%
  mutate(pt = round(fte * (1 / (ptincrease + 0.5)))) %>%
  select(-fte, - ptincrease)

survey_2 <- bind_rows(
  survey_2 %>% filter(ft < 1) %>%
    mutate(ft = 1),
  survey_2 %>% filter(ft >= 1)
)

survey_2
```

# Non-response

A professional mail survey with follow up found a non-response rate of 61% [@nonresponse].
We will also assume that some people will mail in answers and fill in the form incompletely.

```{r}
num_cols_1 = 11

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(nonresponse = sample(c("yes","partial","no"), size = num_total, prob = c(0.6, 0.01, 0.39), replace = TRUE)) %>%
  filter(nonresponse != "no")

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(which_partial = sample(c(4:num_cols_1), size = first(count(survey_1)), replace = TRUE))

for (i in c(1:as.numeric(count(survey_1)))){
  if (survey_1[i, "nonresponse"] == "partial") {
    survey_1[i, as.numeric(survey_1[i,"which_partial"])] <- NA
  }
}

survey_1 <- survey_1 %>%
  select(-nonresponse, -which_partial)

survey_1
```

```{r}
num_cols_2 = 12

set.seed(19894)
survey_2 <- survey_2 %>%
  mutate(nonresponse = sample(c("yes","partial","no"), size = num_total, prob = c(0.6, 0.01, 0.39), replace = TRUE)) %>%
  filter(nonresponse != "no")

set.seed(19894)
survey_2 <- survey_2 %>%
  mutate(which_partial = sample(c(4:num_cols_1), size = first(count(survey_2)), replace = TRUE))

for (i in c(1:as.numeric(count(survey_2)))){
  if (survey_2[i, "nonresponse"] == "partial") {
    survey_2[i, as.numeric(survey_2[i,"which_partial"])] <- NA
  }
}

survey_2 <- survey_2 %>%
  select(-nonresponse, -which_partial)

survey_2
```

```{r}
write_csv(survey_1, here("outputs/data/survey_1.csv"))
write_csv(survey_2, here("outputs/data/survey_2.csv"))
```

\newpage

# References