### 2021-02-23 outputs/paper/paper.Rmd

---
title: "COVID-19 Closures Negatively Impact Restaurant Revenue and Employment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "September 20th, 2021"
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "A random controlled trial, with surveys, to identify the impact of mandatory partial closures on Ontario restaurants"
abstract: "This report covers an experiment conducted by Petit Poll to test the impact of COVID-19 closures on Ontario restaurants. A randomized controlled trial was conducted and surveys were used to collect results. The data show that restaurant revenue and employee counts were negatively impacted by a mandatory two-week partial closure, and dine-in restaurants and restaurants owned by visible minorites were more negatively impacted. These results suggest that if more closures are necessary, Ontario should support restaurant owners by offering support for take-out services and minority-owned businesses."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(finalfit) # for a summary table
library(gridExtra)  # for paired graphs
library(cowplot)  # for plotting side by side
library(lubridate)  # for timeline

# NOTE: scripts 01_scrape_health_depts, 
# 02_stratified_cluster_sampling,
# 03_clean_restaurant_list_csvs, 
# 04_restaurant_survey_frame,
# 05_import_yelp_data,
# and Rmarkdown file data_simulation must have been run already

```

# Introduction

The COVID-19 pandemic has presented challenges that continue to test our resilience, creativity, and ability to recover. From online learning to curbside shopping, the changes are tangible on a local level. To better understand the impact of intermittent lockdowns on restaurant businesses specifically, and to provide the Ontario Ministry of Economic Development, Job Creation and Trade with evidence-based advice, Petit Poll conducted a comprehensive study of over 1,200 restaurants in 12 regions of Ontario. This report describes the experiment design and rationale, measurement strategies, data characteristics, main findings, recommendations, limitations, and broader implications. Importantly, the intervention revealed three key effects of closures: (1) closures had a negative impact on revenues and employment counts, (2) closures had a stronger negative impact on dine-in only restaurants, and (3) closures had a stronger negative impact on Indigenous or visible minority-owned businesses. These findings provide more detail to an already stark picture: while the health of Ontarians has to be prioritized, and the transmission of COVID-19 has to be aggressively contained, it is impossible to ignore the economic consequences of prolonged business closures. In particular, support for Indigenous and visible minority-owned businesses is imperative, as these populations continue to bear the brunt of systemic inequalities. Preparing and supporting dine-in establishments to transition to take-out only, and providing furloughed workers with vital financial support, must be equally prioritized.

# Data

Analysis for this report uses the `R` statistical programming language [@R], and more specifically, the `tidyverse` package for data manipulation [@tidyverse]. To facilitate a reproducible workflow, `here` is used to reference file locations [@here]. Graphs and tables use features from `cowplot` [@cowplot], `finalfit` [@finalfit], `lubridate` [@lubridate], and `kableExtra` [@kableExtra]. Finally, `bookdown` is used to format the report [@bookdown].

In addition to survey data collected by Petit Poll, this report relies on publicly available federal census data [@census_2016] and data from provincial public health departments [@inspection].

This section presents details on the data collection approach and methods, and the steps taken to ensure accuracy.

## Methodology

This experiment used two-stage stratified cluster sampling for survey data collection, a process by which a population is divided into clusters, and these clusters are subsequently divided into strata. A random sample is drawn from each stratum and each cluster. Use of a simple random sample at each stage increases the external validity of the experiment, or the accuracy with which the sampled population represents the target population  [@gertler].

The target population for this study was all restaurants in Ontario. Because restaurant inspection is carried out by different authorities across the province, a list of Ontario local health authorities (LHAs) was collected. These LHAs were the primary sampling units, forming the primary sampling frame, or list of units of interest from which to draw a sample. Using census data from 2016, the primary sampling frame was stratified into small LHAs (population of less than 150,000), medium LHAs (population of 150,000 to 400,000), and large LHAs (population of more than 400,000). Each strata contained the same number of LHAs. From each stratum, two LHAs were randomly sampled to participate in the treatment, and two LHAs were randomly selected to participate in the control. This primary sample is shown in Table \@ref(tab:treatmentcontrolgroups).

This cluster design was used because pandemic shutdowns in Ontario have operated based on LHA, and the goal was to recreate the effect as closely as possible. It also allowed for more consistent enforcement of closures. The stratification was used because there were very differently sized LHAs, the larger typically in urban areas and the smaller in rural areas, and randomly selecting only 12 LHAs from a list of 33 left too high a likelihood of differently sized treatment and control groups. Stratification can create a more balanced survey sample [@samplingtheory]. Given the limitations of sampling from a primary sampling frame of only 33 clusters, stratifying the frame by size of cluster helped ensure the treatment and control groups would be comparable and that the sample population would be representative of the target population.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Import table that shows LHAs selected for treatment and control groups
treatment_control_groups <- read_csv(here::here("outputs/data/treatment_control_groups.csv"))
```

```{r treatmentcontrolgroups, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show LHAs that were selected for treatment and control groups
treatment_control_groups %>%
  knitr::kable(caption = "Primary sample of Local Health Authorities from strata based on population size",
               col.names = c("Group", 
                             "Large Population LHAs", 
                             "Medium Population LHAs", 
                             "Small Population LHAs"),
               booktabs = TRUE) %>%
  kableExtra::kable_styling() %>%
  collapse_rows() %>%
  column_spec(1, color = "#972D15")
```

Once the treatment and control LHAs were selected, each corresponding Food Inspection unit was contacted, and a list of all registered restaurants, the secondary sample units, in each LHA was obtained. Each restaurant was listed by name and address, forming the secondary sampling frame. A portion of this frame is shown in Table \@ref(tab:sampleallunitsdata).

```{r, include = FALSE}
# import table that contains the complete list of restaurants in treatment and control groups
all_units_data <- read_csv(here::here("outputs/data/all_units_data.csv")) 
```

```{r sampleallunitsdata, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show a few rows of the complete list of restaurants in treatment and control groups
all_units_data %>%
  slice(101,1100,2101,3101,4100,5100,6100,7100,8100,9101) %>%
  knitr::kable(caption = "Example segment of the secondary sampling frame",
               booktabs = TRUE) %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "14em") %>%
  column_spec(3, width = "6em") %>%
  column_spec(4, width = "6em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
  row_spec(0, color = "#972D15")
```

A simple random sample of 15% of the treatment list and 15% of the control list was selected to be surveyed, forming the secondary sample. As a result of the simple random sample, the number of restaurants from each LHA was roughly proportionate to the total population of restaurants in each LHA. This selection was used to create a panel, so the same restaurants would be contacted for the first survey and the second survey. Attempting to sample only 15% of the restaurants allowed time and money to be spent on follow-up and multiple methods of data collection, reducing the non-response bias. The randomly sampled restaurants were all assigned ID numbers in a random order.

```{r, include = FALSE}
# import the table that was used to send out survey mailers
table_for_surveys <- read_csv(here::here("outputs/data/table_for_surveys.csv"))
```

The table for surveys was used to generate 2,006 mailers to be sent to each restaurant on the list. Each mailer was a small envelope containing a sheet that invited the restaurant owner to participate, explained the survey, provided a link to the survey, and provided a QR code that went to the same place as the link. There was also a copy of the survey contained in a mailer envelope and a contact number, allowing restaurant owners to complete the survey by phone or mail if they did not feel able to complete it online. If there was no response within two weeks, the restaurant was contacted by phone, and again five days later if there was no response. This procedure was repeated for Survey 1 and Survey 2.

Each link and paper survey was personalized to the restaurant's ID number. The link lead to the survey, which asked for confirmation of the last three digits of the restaurant's postal code. This was to ensure that no mailing mistakes were made, and to prevent duplicate data collection. Once verified, the survey-taker was given the main survey. The survey portion of the experiment cost $6,711.18. A detailed breakdown of costs can be found in [Appendix A](#appendixA).

The data from the survey was tied to the ID number and last three digits of the postal code, but not the restaurant name or address, and the restaurant owner's name was not collected. This was to ensure data privacy. The identifying information was kept in a separate dataframe (Table for Surveys) than the survey answers (Survey 1 Data, Survey 2 Data).

```{r timeline, fig.cap = "Timeline of restaurant closure experiment with surveys", fig.width = 8, fig.height = 4, echo = FALSE, message = FALSE, warning = FALSE}
### from https://benalexkeen.com/creating-a-timeline-graphic-using-r-and-ggplot2/

# create basic dataframe with the milestone events on the line
df <- data.frame(
  year = c(2021, 2021, 2021, 2021, 2021, 2021, 2021),
  month = c(6, 6, 6, 7, 7, 8, 8),
  day = c(3, 26, 28, 1, 15, 3, 27),
  milestone = c("Survey 1 begins", "Survey 1 ends", "Closures \nannounced", "Closures begin", "Closures end", "Survey 2 begins", "Survey 2 ends"),
  status = c("Survey 1", "Survey 1", "Treatment", "Treatment", "Treatment", "Survey 2", "Survey 2")
)

# append event date
df$date <- with(df, ymd(sprintf('%04d%02d%02d', year, month, day)))
df <- df[with(df, order(date)), ]

# set the colors for the milestone events
status_levels <- c("Survey 1", "Treatment", "Survey 2")
status_colors <- c("#D8B70A", "#972D15", "#5BBCD6")

df$status <- factor(df$status, levels=status_levels, ordered=TRUE)

# decide if the milestones will be labelled above or below the line
directions <- c(1, 1, -1, -1, -1, 1, 1)

# set the distance and direction of milestone labels
line_pos <- data.frame(
    "date"=unique(df$date),
    "position"= c(0.2, 0.2, -0.2, -0.4, -0.6, 0.2, 0.2),
    "direction"=rep(directions, length.out=length(unique(df$date)))
)

df <- merge(x=df, y=line_pos, by="date", all = TRUE)
df <- df[with(df, order(date, status)), ]

text_offset <- 0.05

df$month_count <- ave(df$date==df$date, df$date, FUN=cumsum)
df$text_position <- (df$month_count * text_offset * df$direction) + df$position

# set the date range for the timeline
month_date_range <- c(as_date("2021-05-01"), as_date("2021-06-01"), as_date("2021-07-01"), as_date("2021-08-01"), as_date("2021-09-01"))  
month_format <- format(month_date_range, '%b')
month_df <- data.frame(month_date_range, month_format)

# plot the data
ggplot(df,aes(x = date, y = 0, col = status, label = milestone)) +
  labs(col = "") +
  scale_color_manual(values = status_colors, labels = status_levels, drop = FALSE) +
  theme_classic() +
  # Plot horizontal black line for timeline
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  # Plot vertical segment lines for milestones
  geom_segment(data = df[df$month_count == 1,], aes(y = position, yend = 0, xend = date), color = 'black', size = 0.2) +
  # plot colored lines for ongoing events
  geom_segment(aes(x = as_date("2021-05-01"), xend = as_date("2021-05-31"), y = 0, yend = 0), color = "#D8B70A", size = 1) +
  geom_segment(aes(x = as_date("2021-07-01"), xend = as_date("2021-07-31"), y = 0, yend = 0, ), color = "#5BBCD6", size = 1) +
  # Plot scatter points at zero and date
  geom_point(aes(y = 0), size = 3) +
  # Don't show axes, appropriately position legend
  theme(axis.line.y=element_blank(),
                 axis.text.y = element_blank(),
                 axis.title.x = element_blank(),
                 axis.title.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_blank(),
                 axis.ticks.x = element_blank(),
                 axis.line.x = element_blank(),
                 legend.position = "bottom"
                ) +
  # Show text for each month
  geom_text(data = month_df, aes(x = month_date_range, y = -0.05, label = month_format), size = 3, vjust = 0.5, color = 'black') +
  # Show text for each milestone
  geom_text(aes(x = c(as_date("2021-06-03"), as_date("2021-06-26"), as_date("2021-06-25"), as_date("2021-07-01"), as_date("2021-07-15"), as_date("2021-08-03"), as_date("2021-08-27")),
                y = text_position, label = milestone),size=3) +
  # add labels for the ongoing events
  geom_text(aes(x = as_date("2021-05-16"), y = 0.1, label = "Study period of Survey 1"), size = 3, color = "#D8B70A") +
  geom_text(aes(x = as_date("2021-07-16"), y = 0.1, label = "Study period of Survey 2"), size = 3, color = "#5BBCD6")
```

A project timeline can be found in Figure \@ref(fig:timeline). Survey 1 was conducted June 3rd to 27th 2021, and it asked about the month of May 2021. The survey can be found online by clicking here: [2021 Ontario Restaurant Survey 1](https://forms.gle/ZDc5HJBWmJu5snHf9).Survey 1 Data was used to confirm that the two-stage stratified cluster sampling had created treatment and control groups that were roughly equivalent. Survey 1 collected the following data:

* Demographic information
  - Disability status: According to the UN Convention on the Rights of Persons with Disabilities, persons with disabilities are described as having "long-term physical, mental, intellectual or sensory impairments which in interaction with various barriers may hinder their full and effective participation in society on an equal basis with others."
  - Indigenous status: According to the Government of Canada, Indigenous people include "First Nations (North American Indian), Métis or Inuit and/or those who reported Registered or Treaty Indian status, that is registered under the Indian Act of Canada, and/or those who reported membership in a First Nation or Indian band". 
  - Visible minority (non-Indigenous) status: "Visible minority" is defined by the Government of Canada as "persons, other than aboriginal peoples, who are non-Caucasian in race or non-white in colour."
  - Gender identity
* Type of service provided (dine-in, take-out, or both)
* Revenue in May 2021
* Employees
  - Number of full-time employees (30 hours/week or more)
  - Number of part-time employees (less than 30 hours/week)

The demographic information was collected so that disparate impacts on different communities could be identified.
Revenue and number of employees were collected as measures of the restaurant's performance and impact on the local employment levels. These were the primary indicators of interest, due to the Ontario Ministry of Economic Development, Job Creation and Trade's focus on economic stability and employment levels.

After Survey 1 data was collected, on June 28th, shutdowns were announced for the six treatment LHAs. The shutdowns ran from July 1st to July 14th, inclusive. This length of time was considered to be the minimum effective length for a shutdown to stop circulation of the virus in accordance to Public Health Ontario guidelines [@pho]. During the shutdown, all restaurants in the treatment LHAs were officially banned from offering dine-in and patio services. Take-out and delivery were permitted. This type of partial shutdown was selected because it has been the mandated type of shutdown during the Grey level, or lockdown, of the provincial COVID-19 response framework [@restrict].

To date, mandated closure enforcement has included issuing fines from \$750 to $100,000 to businesses found in violation of the province of Ontario's emergency orders [@enforce]. Additionally, provincial offences officers, including police officers, have the authority to disperse crowds indoors as well as outdoors [@police].

Because it is very difficult for a restaurant to move location in under two weeks, the control and treatment groups were effectively separated.

Survey 2 was conducted August 3rd to 27th 2021, and it asked about the month of July 2021. The survey can be found online by clicking here: [2021 Ontario Restaurant Survey 2](https://forms.gle/JWn6VU299uTyufj99). Survey 2 followed a similar format as Survey 1, with the addition of a question about closures. It collected the following data:

* Demographic information
  - Disability status
  - Indigenous status
  - Visible minority (non-Indigenous) status
  - Gender identity
* Type of service provided (dine-in, take-out, or both)
* Closures (none, temporary, or permanent)
* Revenue in July 2021
* Employees
  - Number of full-time employees (30 hours/week or more)
  - Number of part-time employees (less than 30 hours/week)

The scope of this experiment required a structured questionnaire that was easy to distribute, easy to answer, cost effective, reliable, and that provided flexibility of mode of response, including online, over the phone, or on paper. With these features in mind, a survey was an appropriate choice. However, due to the rigidity of pre-determined questions, the survey format might lack potential depth, reducing complex circumstances to simple data points that are easier to collect, manage, and analyze. To account for this, the questions were carefully designed to gather, exactly and only, the necessary data for our study.

[Appendix B](#appendixB) contains screen captures of the online version of Survey 1 and Survey 2.
For images of the sheets included in the mailers sent to restaurant owners, containing all survey questions and answer options for Survey 1 and Survey 2, refer to [Appendix C](#appendixC).

```{r, include = FALSE}
# import survey 1 and 2 data, relevelling the factors so that tables and graphs will display correctly
survey_1 <- read_csv(here("outputs/data/survey_1.csv")) %>%
    mutate(group = fct_relevel(group, "treatment", "control"),
           service_type = fct_relevel(service_type, "dinein", "both", "takeout"),
           disability = fct_relevel(disability, "yes", "no", "nonanswer"),
           woman = fct_relevel(woman, "yes", "no", "nonanswer"),
           indigenous = fct_relevel(indigenous, "yes", "no", "nonanswer"),
           visible = fct_relevel(visible, "yes", "no", "nonanswer"))

survey_2 <- read_csv(here("outputs/data/survey_2.csv")) %>%
    mutate(group = fct_relevel(group, "treatment", "control"),
           service_type = fct_relevel(service_type, "dinein", "both", "takeout"),
           disability = fct_relevel(disability, "yes", "no", "nonanswer"),
           woman = fct_relevel(woman, "yes", "no", "nonanswer"),
           indigenous = fct_relevel(indigenous, "yes", "no", "nonanswer"),
           visible = fct_relevel(visible, "yes", "no", "nonanswer"))

# count the full number of restaurants who replied to at least one survey
count(full_join(survey_1, survey_2, by = "ID"))
```

## Results ^[More information about the Survey 1 and Survey 2 results can be found in [the report *Data for the Ontario Restaurant Closures Experiment*](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment/blob/main/outputs/paper/data_simulation.pdf).]

Reaching restaurant owners to better understand their businesses, and the impact of closures, was a crucial component of this experiment. Approximately 40% of individuals selected to participate completed or partially completed the surveys. This section provides an overview of the data gathered from each survey.

### Survey 1

The initial survey provided a clear description of the characteristics of the treatment and control groups. As indicated in Table \@ref(tab:survey1summary), the groups are comparable, showing no notable differences across any of the variables studied. Survey 1 was necessary to collect data before the intervention to ensure that the treatment and control groups were in fact comparable, verifying the efficacy of the randomized treatment assignment and allowing for an accurate estimation of the counterfactual. The counterfactual – or the representation of what would have happened had the intervention not taken place – is essential to measuring, and reporting on, the effect of the mandated restaurant closures. If the control is an accurate estimate of the counterfactual, the experiment is internally valid and we can observe the true impact of the treatment [@gertler].

The baseline distributions for reported revenue in the treatment and control groups, based on information from May 2021, can be found in Figure \@ref(fig:baselinerevenue). The baseline distributions for number of full-time and part-time employees in the treatment and control groups, also based on information from May 2021, is shown in Figure \@ref(fig:baselineemployees). In this figures, the distributions are extremely similar, indicating the comparability of the treatment and control populations with regard to revenue and employment numbers.

```{r survey1summary, echo = FALSE, message = FALSE, warning = FALSE}
# summarize results for survey 1
survey_1 %>% 
  summary_factorlist("group", c("service_type", "disability", "woman", "indigenous", "visible", "revenue", "ft", "pt"),
                     p = FALSE, 
                     add_dependent_label = TRUE,
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     include_row_missing_col = FALSE,
                     col_totals_rowname = "",
                     total_col = FALSE,
                     col_totals_prefix = "N (%) = ") %>%
  mutate(group = c("", "Service type", "", "", "Owner has disability", "", "", "Owner identifies as woman", "", "", "Owner is Indigenous", "", "", "Owner is a visible minority (non-Indigenous)", "", "", "Revenue", "Number of full-time employees", "Number of part-time employees")) %>%

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control baselines",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
  kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#fbf7e6") %>%
  column_spec(1, color = "#972D15")
```

```{r baselinerevenue, fig.cap="Revenue distribution for treatment and control baselines, from Survey 1", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 1 revenue distribution, separating treatment and control
ggplot(survey_1, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in May 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "Revenue distributions were similar in the treatment and control samples") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

```{r baselineemployees, fig.cap="Employment distribution for treatment and control baselines, from Survey 1", fig.width=10, fig.height=6, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 1 employement numbers, separating treatment and control

# make the full time employment count plot
p1 <- ggplot(survey_1, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125,150)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_1, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees")  +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# using https://stackoverflow.com/questions/37052059/how-to-add-a-title-for-a-grid-layout-figure-in-ggplot2
# display the full-time and part-time plots side by side
p <- plot_grid(p1, p2, align = "h", nrow = 1, rel_widths = c(0.555,0.445))
title <- ggdraw() + draw_label("Employment distributions were similar in the treatment and control samples")
plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
```

### Survey 2

The second survey results showed marked differences between treatment and control groups in most variables. A summary of these results is found in Table \@ref(tab:survey2summary). Naturally, a salient difference is found in service type, with dine-in restaurants decreasing due to mandated shutdowns – the intervention itself – with 7% of restaurants in the control group providing dine-in service, and only 4% in the treatment group. Similarly, the mean revenue decreases noticeably for the treatment group (\$45,959.60) compared to the control group ($70,675.60). The mean numbers of employees, both full-time and part-time, show a decrease as well, from 8 and 5 respectively in the control group, to 5 and 3 for the treatment. A visual representation of the differences between treatment and control groups for revenue and number of employees can be found in Figure \@ref(fig:postrevenue) and Figure \@ref(fig:postemployees) respectively.

```{r survey2summary, echo = FALSE, message = FALSE, warning = FALSE}
# summarize results for survey 1
survey_2 %>% 
  summary_factorlist("group", c("service_type", "disability", "woman", "indigenous", "visible", "shutdown", "revenue", "ft", "pt"),
                     p = FALSE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     include_row_missing_col = FALSE,
                     col_totals_rowname = "",
                     total_col = FALSE,
                     col_totals_prefix = "N (%) = ") %>%
  mutate(group = c("", "Service type", "", "", "Owner has disability", "", "", "Owner identifies as woman", "", "", "Owner is Indigenous", "", "", "Owner is a visible minority (non-Indigenous)", "", "", "Closure", "", "", "Revenue", "Number of full-time employees", "Number of part-time employees")) %>%

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control groups post-treatment",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
   kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#fbf7e6") %>%
  column_spec(1, color = "#972D15")
```

# Discussion

Petit Poll commends the Ontario Ministry of Economic Development, Job Creation and Trade for its commitment to learn more about the impact of restaurant closures in the province. This study has uncovered three main findings, supplemented in this section by a series of recommendations. Additionally, this section elaborates on relevant ethical considerations, potential biases and limitations. The final section presents some initial thoughts for future related research.

## Overview

Understanding the impact of COVID-19 closures on small businesses can help the Government of Ontario prepare to rebuild once the immediate health threat is contained. In particular, the contributions of restaurant businesses to the local economy have to be studied in terms of employment and revenue generation. According to The Associated Press, there was a 20% decline in employment in the restaurant, hotel, and entertainment sector between November 2019 and November 2020 [@ap]. The National Restaurant Association reports a total loss industry-wide of $120 billion between March and May 2020 alone [@rd]. Although these figures are based on United States data, the findings from our experiment show a similarly concerning reality in Ontario.

The experiment, which consisted of an initial survey, a two-week restaurant closure intervention, and a second survey, revealed that the impact of closures is not distributed equally across service types, nor across restaurant owner demographics. Swift action to protect and uplift the local economy, particularly dine-in restaurants and restaurants owned by Indigenous peoples or members of visible minority groups, will be crucial in the upcoming months.

## Findings

To prevent a misleading estimate of the counterfactual in this experiment, the effects of the intervention are not evaluated by comparing the survey results before and after the intervention, but rather by comparing the treatment and control groups after the intervention. The following subsections provide further detail on each of the main findings.

### Closures had a negative impact on revenues and employment counts

As noted in the results from Survey 2, the intervention led to a noticeable difference in reported revenues. The mean reported revenue for the treatment group was \$45,959.60 and $70,675.60 for the control. The revenue distribution for both groups is show in Figure \@ref(fig:postrevenue). The mean number of employees, both full-time and part-time, show a decrease as well, from 8 and 5 respectively in the control group, to 5 and 3 for the treatment. Figure \@ref(fig:postemployees) shows further detail of this finding.

With these effects in mind, Petit Poll supports the Government of Ontario's programs in benefit of small businesses: the Ontario Small Business Support Grant, the Personal Protective Equipment Relief Grant, and the current Property Tax and Energy Rebate Grants [@relief]. Additionally, based on the evidence shown in this report, it is important to consider developing provincial financial support programs for furloughed employees, in addition to standing federal programs like Employment Insurance, that directly address the negative impact of COVID-19 closures. Ensuring the protection of workers in the restaurant industry specifically, not only physically as the spread of the virus is contained, but economically with financial aid will be crucial to the success of mandated closures.

```{r postrevenue, fig.cap="Revenue distribution for treatment and control groups, from Survey 2", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 2 revenue distribution, separating treatment and control
ggplot(survey_2, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in July 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "The revenue distribution was lower for the treatment sample") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

```{r postemployees, fig.cap="Employment distribution for treatment and control groups, from Survey 2", fig.width=10, fig.height=6, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 2 employement numbers, separating treatment and control

# make the full time employment count plot
p1 <- ggplot(survey_2, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_2, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# using https://stackoverflow.com/questions/37052059/how-to-add-a-title-for-a-grid-layout-figure-in-ggplot2
# display the full-time and part-time plots side by side
p <- plot_grid(p1, p2, align = "h", nrow = 1, rel_widths = c(0.555,0.445))
title <- ggdraw() + draw_label("Employment distributions were lower for the treatment sample")
plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
```

### Closures had a stronger negative impact on dine-in only restaurants

It is worth noting that the negative impact on revenue was not equally distributed across the three types of establishments. Compared to the control group, treatment group dine-in restaurants suffered significant losses in average revenue. Restaurant owners in the treatment group who only offered dine-in service reported a mean revenue of \$34,609.40, those who offered both take-out and dine-in service reported \$42,889.41, and those who offered only take-out service reported \$53,156.02. The control group reported \$67,744.79, \$70,062.17, and \$72,860.20 for dine-in only, both, and take-out only respectively. Figure \@ref(fig:servicetypeimpact) shows a comparison of revenue distributions for restaurants offering dine-in service, take-out, or both.

The effect of mandated closures is attenuated for take-out only restaurants. This indicates that providing restaurant owners with support to transition their regular operations to take-out only service should be considered. The City of Toronto-funded Digital Main Street, created by the Toronto Association of Business Improvement Areas, might serve as an exemplary approach. The program provides businesses with the tools and information to build an online presence, accommodating for physically-distanced shopping. The free program includes an online learning platform, training programs, and dedicated support staff [@mainstreet]. A parallel provincial program aimed directly at restaurant owners who wish to transition out of dine-in only service might help alleviate the economic pressures of COVID-19 mandated closures, while maintaining the necessary health emergency protocols.

```{r servicetypeimpact, fig.cap = "Survey 2 revenue distributions for service types with means marked", fig.width=7, fig.height=5, echo = FALSE, message = FALSE, warning = FALSE, out.width = '80%', fig.align = "center"}
# plot distributions of revenue, split by treatment/control and by type of service offered

# from https://stackoverflow.com/questions/47651868/split-violin-plot-with-ggplot2-with-quantiles
GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, draw_group = function(self, data, ..., draw_quantiles = NULL){
  data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
  grp <- data[1,'group']
  newdata <- plyr::arrange(transform(data, x = if(grp%%2==1) xminv else xmaxv), if(grp%%2==1) y else -y)
  newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
  newdata[c(1,nrow(newdata)-1,nrow(newdata)), 'x'] <- round(newdata[1, 'x']) 
  if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
    stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <= 
                                              1))
    quantiles <- create_quantile_segment_frame(data, draw_quantiles)
    aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
    aesthetics$alpha <- rep(1, nrow(quantiles))
    both <- cbind(quantiles, aesthetics)
    quantile_grob <- GeomPath$draw_panel(both, ...)
    ggplot2:::ggname("geom_split_violin", grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
  }
  else {
    ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
  }
})
geom_split_violin <- function (mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, position = position, show.legend = show.legend, inherit.aes = inherit.aes, params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
}

# create summary data of means for the plot
means <- survey_2 %>%
  filter(service_type != "NA" & revenue != "NA") %>%
  group_by(group, service_type) %>%
  summarise(mean = mean(revenue) / 1000)

# create plot using summary data of means, original dataframe, and the split violin function above:
ggplot() +
  geom_split_violin(data = (survey_2 %>% filter(service_type != "NA")), aes(x = service_type, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means, 
                aes(x = service_type, y = mean, ymax = mean, ymin = mean, fill = group), 
                position = "dodge",
                color = "#972D15") +
  labs(fill = NULL,
       x = "Service type provided",
       title = "Revenues were more heavily impacted for dine-in only establishments",
       y = "Revenue (in thousands of Canadian dollars)") +
  scale_x_discrete(labels = c( "Dine-in only", "Both dine-in and take-out", "Take-out only")) +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

### Closures had a stronger negative impact on Indigenous or visible minority-owned businesses

The negative impact of mandated closures was not distributed equally across restaurant owner demographics either. As indicated in Figure \@ref(fig:minorityimpact), treatment group restaurant owners who identified as Indigenous or as members of a visible minority group reported a larger loss in average revenues, compared to the control, than those who did not identify as Indigenous or a visible minority. The mean revenue for those who identified as visible minority or Indigenous in the treatment group was \$28,749.17, and it was \$47,572.93 for those who did not. In the control group, the mean revenue for those who identified as Indigenous or as members of a visible minority group was \$73,983.60, and it was \$70,777.57 for those who did not.

Protecting populations that face systemic inequalities is absolutely imperative to the success of any proposed pandemic response. Following the approach of the federal government might be a fruitful start. Currently, the Government of Canada offers interest-free loans and non-repayable contributions to Indigenous-owned businesses [@ibi]. Further, the federal government has announced and collected applications to the Black Entrepreneurship Program, an initiative to ensure Black-owned businesses thrive [@bep]. Petit Poll recommends Ontario take similar, yet bolder, steps to ensure Indigenous and visible minority owners are not left behind during the COVID-19 pandemic. To do so, restaurant industry-specific interest-free loans, non-repayable contributions, free educational programming, mentorship, and additional supports should be extended to Indigenous and visible minority restaurant owners across the province.

```{r minorityimpact, fig.cap="Survey 2 revenue distributions for minority status with means marked", fig.width=6.25, fig.height=5, echo = FALSE, message = FALSE, warning = FALSE, out.width = '70%', fig.align = "center"}
# plot distributions of revenue, split by treatment/control and by type of service offered

# create summary data for the plot
data <- bind_rows(
  survey_2 %>%
    filter(indigenous == "yes" | visible == "yes") %>%
     mutate(minority = "Visible minority / Indigenous")
  ,
  survey_2 %>%
   filter(indigenous == "no" & visible == "no") %>%
    mutate(minority = "No")
) %>% 
  mutate(minority = fct_relevel(minority, "Visible minority / Indigenous", "No"))

# create summary data of means for the plot
means <- data %>%
  filter(service_type != "NA" & revenue != "NA") %>%
  group_by(group, minority) %>%
  summarise(mean = mean(revenue) / 1000)

# create plot using summary data of means, original dataframe, and the split violin function above:  
ggplot() +
  geom_split_violin(data = data, aes(x = minority, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means,
                aes(x = minority, y = mean, ymax = mean, ymin = mean, fill = group),
                position = "dodge",
                color = "#972D15") +
  labs(fill = NULL,
       x = "Racial and ethnic minorities",
       title = "Revenues were more heavily impacted for minority-owned businesses",
       y = "Revenue (in thousands of Canadian dollars)") +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

### Implications

The findings from this experiment uncover the negative impact that mandated closures have on restaurant businesses specifically. In a national emergency, these effects have to be contextualized by assessing the impact on a broader societal level. As the COVID-19 pandemic continues to test our resilience, it is important to consider the economic inter-dependencies that rely on the health of restaurant businesses. As noted in Forbes [-@forbes], a key feature of neighbourhood desirability is access to a wide range of shops and restaurants. Restaurant closures, leading to vacant commercial lots, might have a negative impact on the real estate market, setting off an unfortunate ripple effect that could destabilize the economy further. Preventing the permanent closure of restaurants might prove to be a worthwhile investment in the months following the pandemic.

On an individual level, these findings point at the importance of employment. Gainful employment is directly tied to psychological wellbeing, with experts citing work as one of the most important determinants of quality of life, second only to personal relationships. [@sage] During a crisis that has the potential to severely impact the population's mental health, ensuring individuals retain their employment is vital. Employment, aside from providing an income and the possibility of financial independence, fosters a positive network of social support — a crucial social determinant of health [@who]. Prioritizing efforts to establish positive networks of social support seems particularly necessary during a pandemic that requires measures of physical distancing.

These examples provide a mere glance at the economic and social implications of our findings. To better understand these implications on a micro and macro level, Petit Poll recommends the Ontario Ministry of Economic Development, Job Creation and Trade, consult with relevant government authorities and departments.

## Ethics

This experiment would not have been possible without the participation of hundreds of restaurant owners. It is Petit Poll's greatest priority to ensure all studies adhere to the highest ethical standards, especially when involving human participants. Ethical requirements for this project, particularly the survey portion, included:

- Anonymity, by determining that collecting the restaurant owner's name was unnecessary, and processing survey data with numerical IDs rather than the restaurant name or address.
- Confidentiality, by employing a professional data management account throughout the entirety of the project.
- Informed consent, by providing a clear explanation of the purposes of the study, the nature of the survey, and contact details should participants require additional information. 
- Voluntary participation and disclosure, by allowing respondents to engage or refuse to engage in the study and allowing non-response of demographic questions.

Additionally, in order to reduce potential biases, the survey portion of this experiment used a thoroughly-tested questionnaire, a robust methodology, expert interviewers, and invested in following up with selected participants who did not initially respond.

Finally, in order to ascertain that the recommendations for each of the findings in this report promote and advance human rights, we highlight the right to work as noted by the Ontario Human Rights Commission (OHRC). In response to the COVID-19 pandemic, the OHRC calls for the upholding of this right by providing special supports to vulnerable groups, including racialized communities and workers in precarious employment. The OHRC recognizes the loss of employment that leads to loss of household income and increased poverty as a direct impact of COVID-19 on vulnerable groups [@ohrc2].

## Limitations

Although great care was taken with every step of the design of this experiment, with the goal of gathering data that would accurately represent current conditions, the authors acknowledge to following limitations.

Cluster sampling has inherent constraints that might affect external validity, including the reduction of the population, in this case the entirety of restaurant business in Ontario, to smaller groups, potentially affecting representation [@roy]. This may lead to a higher potential of sampling error and biases, as the factor by which the population is clustered might introduce unforeseen conditions [@roy]. However, cluster sampling by size of local health authority population was necessary as the goal of this experiment was to replicate current protocols for mandated closures, which have been dictated by local health authorities. The ethical implications of experimenting in the entirety of the population proved to be unsound, as the negative effects of closures would have been exacerbated. Sampling was necessary to understand the impact while containing its potentially devastating effects.

A second limitation arises from the decision to use food inspection records to collect information on current restaurants. We acknowledge that this process would effectively erase from our study all businesses that do not currently operate in accordance to the legal requirement to properly license a restaurant business. As such, should the Ontario Ministry of Economic Development, Job Creation and Trade need to understand the COVID-19 impact on non-licensed restaurant businesses, we recommend further research.

Two types of survey non-response, unit and item non-response, present another limitation. The first, in which an individual refuses to participate, and the second, in which a participant refuses to answer a particular question, might affect the accuracy of inferences if the individuals who participate are systemically different from those who do not [@imai]. This study reduced the impact of non-response bias by allowing participants to complete the survey by online form, mail, or phone, and by following up on non-responses.

```{r nonresponse, fig.cap="Non-response and partial response rates for Survey 1 and 2", fig.width=8, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot nonresponse, response, and partial response numbers for survey 1 and survey 2

# calculate relevant numbers for the plot for the sample
num_treat_sample <- as.numeric(count(table_for_surveys %>% filter(group == "treatment")))

num_control_sample <- as.numeric(count(table_for_surveys %>% filter(group == "control")))

# calculate relevant numbers for the plot for survey 1
num_treat_complete_survey1 <- sum(complete.cases(survey_1 %>% filter(group == "treatment")))

num_control_complete_survey1 <- sum(complete.cases(survey_1 %>% filter(group == "control")))

num_treat_partial_survey1 <- sum(!complete.cases(survey_1 %>% filter(group == "treatment")))

num_control_partial_survey1 <- sum(!complete.cases(survey_1 %>% filter(group == "control")))

num_treat_nonre_survey1 <- num_treat_sample - num_treat_complete_survey1 - num_treat_partial_survey1

num_control_nonresurvey1 <- num_control_sample - num_control_complete_survey1 - num_control_partial_survey1

# calculate relevant numbers for the plot for survey 2
num_treat_complete_survey2 <- sum(complete.cases(survey_2 %>% filter(group == "treatment")))

num_control_complete_survey2 <- sum(complete.cases(survey_2 %>% filter(group == "control")))

num_treat_partial_survey2 <- sum(!complete.cases(survey_2 %>% filter(group == "treatment")))

num_control_partial_survey2 <- sum(!complete.cases(survey_2 %>% filter(group == "control")))

num_treat_nonre_survey2 <- num_treat_sample - num_treat_complete_survey2 - num_treat_partial_survey2

num_control_nonresurvey2 <- num_control_sample - num_control_complete_survey2 - num_control_partial_survey2

# make tibble containing above numbers
tibble(Group = c("Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control"), 
        Response = c("Complete", "Complete", "Partial", "Partial", "No response", "No response", "Complete", "Complete", "Partial", "Partial", "No response", "No response"), 
        Survey = c("Survey 1", "Survey 1","Survey 1","Survey 1","Survey 1","Survey 1", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2"), 
        Count = c(num_treat_complete_survey1, num_control_complete_survey1, num_treat_partial_survey1, num_control_partial_survey1, num_treat_nonre_survey1, num_control_nonresurvey1,
          num_treat_complete_survey2, num_control_complete_survey2, num_treat_partial_survey2, num_control_partial_survey2, num_treat_nonre_survey2, num_control_nonresurvey2)) %>%
  # plot the data
  ggplot(aes(x = Response, y = Count, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "More than half of surveyed restaurants did not respond") +
  facet_wrap(~Survey) +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

A final limitation is the use of self-reported information. A main concern is the incidence of social desirability bias, in which respondents might provide answers they deem socially acceptable, rather than the truth [@imai]. In this experiment, such bias could be present in the answers of Survey 2 which asks respondents about closures. Victims to this type of bias, respondents might have claimed they voluntarily ceased their restaurant operations to limit the spread of COVID-19 even if they had not.

## Future Directions

For extrapolating results to the whole of Ontario, the demographics of the areas sampled should be compared to the demographics of the entire province. While efforts were taken to take a sample of LHA areas representative of all those in Ontario, extreme demographic variation between LHAs made this difficult. Considering the LHAs selected for the treatment group, the population is 3.8% Indigenous and 11.9% visible minority [@census_2016]. If we look at the LHAs selected for the control group, the population is 4.7% Indigenous and 16.6% visible minority [@census_2016]. Ontario's population as a whole is 2.8% Indigenous and 28.9% visible minority [@census_2016]. While the units of study in this experiment were restaurants and not individuals, it is worth considering that there may proportionately be more minority-owned restaurants in the entire province than these surveys suggest. As minority-owned restaurants were more impacted by the closures in this study, the proportion of minority-owned restaurants may influence the outcome of a pandemic closure. Therefore, this experiment should be considered a first step in understanding the impact of COVID-19 closures and additional research, aimed deliberately at understanding minority-owned businesses, should be conducted.

This experiment collected information about restaurant owners in Ontario to study the economic and social impact of closures on a business level. To understand the effects on an individual level, questionnaires that are aimed at employees would be a feasible adaptation of this study.

Particularly, future research might consider studying the impact of closures on populations made vulnerable by the COVID-19 pandemic. The discrimination against the East Asian community that the pandemic exacerbated [@ohrc1], fueled by sensationalist media reports and mis- and disinformation online, might have targeted restaurant businesses directly. A comprehensive study would reveal further information.

Finally, this experiment did not differentiate between independent restaurant businesses and franchises or chains. To better develop support programs, it might be advantageous to deliberately inquire about the type of business model under which the restaurant operates. As this experiment revealed, dine-in and visible minority-owned restaurants were the most affected by closures. Understanding whether take-out service is more frequent in franchises, or if less members of visible minority groups own independent restaurants, for example, would allow for the informed allocation of funds towards targeted support programs.

\newpage

# Appendix A {#appendixA}

```{r echo=FALSE, message=FALSE, warning=FALSE}
# import survey budget table
survey_budget <- read_csv(here("outputs/data/survey_budget.csv"))

# display survey budget in a table
survey_budget %>% 
  kable(digits = 2,
        caption = "2021 Ontario Restaurant Survey Budget") %>% 
  kable_styling() %>% 
  row_spec(0, bold = T, color = "white", background = "#972D15") %>% 
  row_spec(4, bold = T) %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```

\newpage

# Appendix B {#appendixB}

```{r page1, echo=FALSE, fig.cap="Screen capture of the first section of Survey 1 and Survey 2.", out.width = '70%', fig.align="center"}
knitr::include_graphics(here("outputs/images/page1.png"))
```

```{r page2a, echo=FALSE, fig.cap="Screen capture of the second section of Survey 1.", out.width = '75%', fig.align="center"}
knitr::include_graphics(here("outputs/images/page2a.png"))
```

```{r page2b, echo=FALSE, fig.cap="Screen capture of the second section of Survey 2.", out.width = '75%', fig.align="center"}
knitr::include_graphics(here("outputs/images/page2b.png"))
```

```{r page3, echo=FALSE, fig.cap="Screen capture of the third section of Survey 1 and Survey 2.", out.width = '75%', fig.align="center"}
knitr::include_graphics(here("outputs/images/page3.png"))
```

```{r page4, echo=FALSE, fig.cap="Screen capture of the fourth section of Survey 1 and Survey 2.", out.width = '75%', fig.align="center"}
knitr::include_graphics(here("outputs/images/page4.png"))
```

\clearpage

# Appendix C {#appendixC}

```{r survey1png, echo=FALSE, fig.cap="Image of the mailer sent to restaurant owners containing Survey 1.", out.width = '78%', fig.align="center"}
knitr::include_graphics(here("outputs/images/survey1.png"))
```

```{r survey2png, echo=FALSE, fig.cap="Image of the mailer sent to restaurant owners containing Survey 2.", out.width = '78%', fig.align="center"}
knitr::include_graphics(here("outputs/images/survey2.png"))
```

\newpage

# References