### 2021-02-25 outputs/paper/data_simulation.Rmd
refine data_simulation document. change title to something more specific.

---
title: "Data for the Ontario Restaurant Closures Experiment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Simulation Methods and Parameterization Research"
abstract: "This document explains how data for the Ontario restaurant closures experiment was simulated, including the research that was used for parameterization and the methods used to randomize data."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in projects

# read in the data
# NOTE: scripts 01_scrape_health_depts, 
# 02_stratified_cluster_sampling,
# 03_clean_restaurant_list_csvs, 
# 04_restaurant_survey_frame,
# and 05_import_yelp_data must have been run already
```

# Introduction

Analysis for this project uses the R statistical programming language [@R], and more specifically, the `tidyverse` package for data manipulation [@tidyverse]. Because the data is managed using R Projects, `here` is used to reference file locations [@here]. `kableExtra` formats tables [@kableExtra]. `bookdown` is used to format the report [@bookdown].

First, we read in the skeleton for the survey data in Table \@ref(tab:start). This is what we would hypothetically know about a restaurant before they took the survey. We keep only the ID numbers, the group, and the last 3 characters of the postal code, which the survey-taker will have to enter to verify identity. All other variables, like name and full address, are only kept in the only frame for surveying, which contains no private information.

```{r start, echo = FALSE, message = FALSE, warning = FALSE}
table_for_surveys <- read_csv(here::here("outputs/data/table_for_surveys.csv"))

survey_1 <- table_for_surveys %>%
  select(ID, group, verify) %>%
  arrange(ID)

# count the numbers for treat, control, and total
num_treat <- first(count(survey_1 %>%
                           filter(group == "treatment")))

num_control <- first(count(survey_1 %>% 
                             filter(group == "control")))

num_total <- first(count(survey_1))

survey_2 <- survey_1

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Frame for surveys") %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

From there, data was simulated for all questions asked on the survey.

# Type of service provided

## Research

First, we used Yelp restaurant data to see how many Ontario restaurants offered different types of service in Table \@ref(tab:yelptakeout) and Table \@ref(tab:yelptable) [@yelp_dataset].

```{r, include = FALSE}
# Read in data
yelp_data <- read_csv(here("inputs/data/yelp_restaurants_ontario.csv")) %>% 
  na.omit() # Remove all NAs
total_restaurants <- count(yelp_data)

# Check proportion of RestaurantsTakeOut
takeout <- yelp_data %>% 
  select(RestaurantsTakeOut)
summary(takeout)
```

```{r yelptakeout, echo = FALSE, message = FALSE, warning = FALSE}
takeout_percentage <- takeout %>%
  group_by(RestaurantsTakeOut) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

takeout_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Yelp Ontario take-out data") %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

```{r, include = FALSE}
# Check proportion of RestaurantsTableService
tableserv <- yelp_data %>% 
  select(RestaurantsTableService)
summary(tableserv)
```

```{r yelptable, echo = FALSE, message = FALSE, warning = FALSE}
tableserv_percentage <- tableserv %>%
  group_by(RestaurantsTableService) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

tableserv_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Yelp Ontario table service data") %>%
  kableExtra::kable_styling() 
```

We used these percentages to estimate proportions of restaurants with take-out and dine-in service, in Table \@ref(tab:serviceprop).

```{r serviceprop, echo = FALSE, message = FALSE, warning = FALSE}
# Calculate invalid responses (neither dine-in nor takeout)
invalid <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == FALSE) %>% 
  count()

valid_restaurants <- total_restaurants-invalid

# Calculate service type percentages
takeout_only <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

dinein_only <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == FALSE) %>% 
  count()/valid_restaurants

both <- yelp_data %>% 
  filter(RestaurantsTableService == TRUE & RestaurantsTakeOut == TRUE) %>% 
  count()/valid_restaurants

service_percentage<- data.frame(
  service = c("takeout", "dinein", "both"),
  percentage = c(takeout_only$n, dinein_only$n, both$n), options(digits = 2)
)

 service_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Estimated service type proportions") %>%
  kableExtra::kable_styling()
```

According to Ontario restaurant data from Yelp [@yelp_dataset], approximately 6% of restaurants
offer dine in service, 32.5% offer takeout, and 61.5% offer both.

## Survey 1 and 2

We randomly sampled to assign each restaurant to be "dinein", "both", or "takeout", with probabilities of 0.06, 0.615, and 0.325, in Table \@ref(tab:servicetype1) and Table \@ref(tab:servicetype2).

```{r servicetype1, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign service_provided using ratios above

set.seed(19)
survey_1 <- survey_1 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding service type to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r servicetype2, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign service_provided using ratios above

set.seed(19)
survey_2 <- survey_2 %>%
  mutate(service_type = sample(x = c("dinein",
                                     "both",
                                     "takeout"),
      size = num_total,
      replace = TRUE, 
      prob = c(0.06, 0.615, 0.325)
      ))

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding service type to Survey 2 data") %>%
  kableExtra::kable_styling()
```

# Demographic traits of owners: disability

## Research

0.5% of small to medium enterprises in Canada are owned by persons with disabilities [@ownershipdemos].

## Survey 1 and 2

We randomly sampled to assign each restaurant to either "yes", "no", or "nonanswer", with probabilities of 0.005, 0.985, and 0.01, in Table \@ref(tab:disability1) and Table \@ref(tab:disability2).

```{r disability1, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign disability status using ratios above

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(disability = sample(c("yes", "no", "nonanswer"),
                             size = num_total,
                             prob = c(0.005, 0.985, 0.01),
                             replace = TRUE))

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner disability status to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r disability2, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign disability status using ratios above

set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(disability = sample(c("yes", "no", "nonanswer"),
                             size = num_total,
                             prob = c(0.005, .985, 0.01),
                             replace = TRUE))

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner disability status to Survey 2 data") %>%
  kableExtra::kable_styling()
```

# Demographic traits of owners: racial and ethnic minorities and gender

## Research

We researched racial and ethnic minorities and women using census data broken down by local health authorities [@census_2016], shown in Table \@ref(tab:LHApop), Table \@ref(tab:groupscount), and Table \@ref(tab:groupsprop).

```{r LHApop, echo = FALSE, message = FALSE, warning = FALSE}
### Load census dataset by Health Regions ###
### Download link: 
### https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/
### download-telecharger/comp/GetFile.cfm?Lang=E&FILETYPE=CSV&GEONO=058
census_2016 <- read.csv(here("inputs", "data", "census_2016.csv"))
census_2016 <- janitor::clean_names(census_2016)

# Narrowed down categories
demographic_info <- c(1, 8, 1324, 1290, 1917)

### Function to get data from specific regions and rows ###

get_region_data <- function(y){
  census_2016 %>%
    filter(geo_code_por == y,
           member_id_profile_of_health_regions_2247 %in% demographic_info) %>%
    select(dim_profile_of_health_regions_2247, 
           member_id_profile_of_health_regions_2247, 
           dim_sex_3_member_id_1_total_sex, 
           dim_sex_3_member_id_2_male, 
           dim_sex_3_member_id_3_female)
}

### Get data from the randomly selected Health Regions ###
# GEO_CODE indicates public health regions, taken from Census 2016 links:
# https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/
# search-recherche/lst/results-resultats.cfm?Lang=E&TABID=1&G=1&Geo1=&Code1=
# &Geo2=&Code2=&GEOCODE=35&type=0

ontario <- get_region_data(35)

haliburton <- get_region_data(3535)
hamilton <- get_region_data(3537)
algoma <- get_region_data(3526)
simcoe_muskoka <- get_region_data(3560)
timiskaming <- get_region_data(3563)
windsor_essex <- get_region_data(3568)

northwest <- get_region_data(3549)
# Southwestern was created by almagamating oxford and elgin-st. thomas units:
oxford <- get_region_data(3552)
elgin <- get_region_data(3531)
waterloo <- get_region_data(3565)
durham <- get_region_data(3530)
sudbury <- get_region_data(3561)
brant <- get_region_data(3527)

### Put together demographic info in one table ###

populationss <- c("Total_Population", 
                  "Indigenous_Population_25%_sample", 
                  "Visible_Minority_25%_sample", 
                  "Accommodation_and_Food_Services_25%_sample", 
                  "Total_Population_Women")

get_pop_info <- function(x){
  c(as.numeric(x$dim_sex_3_member_id_1_total_sex
               [x$dim_profile_of_health_regions_2247 == "Population, 2016"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex
               [x$dim_profile_of_health_regions_2247 == "Aboriginal identity"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex
               [x$dim_profile_of_health_regions_2247 == 
                   "Total visible minority population"]),
    as.numeric(x$dim_sex_3_member_id_1_total_sex
               [x$dim_profile_of_health_regions_2247 == 
                   "72 Accommodation and food services"]),
    as.numeric(x$dim_sex_3_member_id_3_female
               [x$dim_profile_of_health_regions_2247 == 
                   "Total - Age groups and average age of the population - 100% data"]))
}

ontario_pop <- get_pop_info(ontario)

haliburton_pop <- get_pop_info(haliburton)
algoma_pop <- get_pop_info(algoma)
hamilton_pop <- get_pop_info(hamilton)
windsor_essex_pop <- get_pop_info(windsor_essex)
simcoe_muskoka_pop <- get_pop_info(simcoe_muskoka)
timiskaming_pop <- get_pop_info(timiskaming)

brant_pop <- get_pop_info(brant)
sudbury_pop <- get_pop_info(sudbury)
#southwestern_pop <- get_pop_info(southwestern)
oxford_pop <- get_pop_info(oxford)
elgin_pop <- get_pop_info(elgin)
northwest_pop <- get_pop_info(northwest)
waterloo_pop <- get_pop_info(waterloo)
durham_pop <- get_pop_info(durham)

populations <- bind_cols(populationss, 
                         ontario_pop, 
                         haliburton_pop, 
                         algoma_pop, 
                         hamilton_pop, 
                         windsor_essex_pop, 
                         simcoe_muskoka_pop, 
                         timiskaming_pop, 
                         brant_pop, 
                         sudbury_pop, 
                         oxford_pop, 
                         elgin_pop, 
                         northwest_pop, 
                         waterloo_pop, 
                         durham_pop)

colnames(populations) <- c("Info", "Ontario", "haliburton", "algoma", 
                           "hamilton", "windsor", "simcoe", 
                           "timiskaming", "brant", "sudbury", "oxford", 
                           "elgin", "northwestern", "waterloo", "durham")

populations <- populations %>%
  mutate(southwestern = oxford + elgin) %>%
  mutate(total_treat = 
           haliburton + algoma + hamilton + windsor + simcoe + timiskaming,
         total_control = 
           brant + sudbury + northwestern + waterloo + durham + southwestern) %>%
  select(-oxford, -elgin)

populations %>%
  slice(1,2,3,5) %>%
  select(-Ontario, -total_treat, -total_control) %>%
  mutate(Info = c("Total Population", "Indigenous", "Visible Minority", "Women")) %>%
  knitr::kable(booktabs = TRUE, caption = "Populations in Ontario local health authorities") %>%
  kableExtra::kable_styling(latex_options = "scale_down")

#save this table
write_csv(populations, here("outputs/data/populations"))
```

```{r groupscount, echo = FALSE, message = FALSE, warning = FALSE}
# Split the population info into our treatment and control groups.

populations_split <- matrix(ncol=4, nrow=5)
populations_split[,1] <- c("Total_Population", 
                           "Indigenous_Population_25%_sample", 
                           "Visible_Minority_25%_sample", "
                           Accommodation_and_Food_Services_25%_sample", 
                           "Total_Population_Women")
populations_split[,2] <- ontario_pop

total_pop_treatment <- sum(populations[1,c(3:8)])
indigenous_treatment <- sum(populations[2,c(3:8)])
minority_treatment <- sum(populations[3,c(3:8)])
food_services_treatment <- sum(populations[4,c(3:8)])
women_treatment <- sum(populations[5,c(3:8)])

treatment <- c(total_pop_treatment, 
               indigenous_treatment, 
               minority_treatment, 
               food_services_treatment, 
               women_treatment)
populations_split[,3] <- treatment

total_pop_control <- sum(populations[1,c(9:14)])
indigenous_control <- sum(populations[2,c(9:14)])
minority_control <- sum(populations[3,c(9:14)])
food_services_control <- sum(populations[4,c(9:14)])
women_control <- sum(populations[5,c(9:14)])

control <- c(total_pop_control, 
             indigenous_control, 
             minority_control, 
             food_services_control, 
             women_control)
populations_split[,4] <- control

populations_split <- as.data.frame(populations_split)
colnames(populations_split) <- c("Info", "Ontario", "Treatment", "Control")

populations_split %>%
  slice(1,2,3,5) %>%
  knitr::kable(booktabs = TRUE, caption = "Populations in treatment and control groups") %>%
  kableExtra::kable_styling()
```

```{r groupsprop, echo = FALSE, message = FALSE, warning = FALSE}
### Get percentage proportions of demographic groups of interest

populations_split_percentage <- matrix(ncol=4, nrow=5)

for(i in 2:length(populations_split)){
  for(j in 1:5){
    populations_split_percentage[j,i] <- 
      round(as.numeric(populations_split[j,i])/as.numeric(populations_split[1,i]), 
            digits=3)
  }
}

#populations_percentage[1, 2:4] <- c(1, 1, 1)
populations_split_percentage[,1] <- c("Total_Population", 
                                      "Indigenous_Population_25%_sample", 
                                      "Visible_Minority_25%_sample", 
                                      "Accommodation_and_Food_Services_25%_sample", 
                                      "Total_Population_Women")

populations_split_percentage <- as.data.frame(populations_split_percentage)
colnames(populations_split_percentage) <- c("Info", 
                                            "Ontario", 
                                            "Treatment", 
                                            "Control")

populations_split_percentage %>%
  slice(1,2,3,5) %>%
  knitr::kable(booktabs = TRUE, caption = "Proportions in treatment and control groups") %>%
  kableExtra::kable_styling()
```

We established expected percentages for the treatment and control regions based on census data [@census_2016].

15.6% of business owners are women in Canada [@ownershipdemos].

Visible minorities are 25% of the population [@visibleminority].

12.2% of business owners are visible minorities [@ownershipdemos].

Therefore, visible minorities are 0.49 as likely to own a business.
Therefore, we can estimate that 0.058 of restaurant owners are a visible minority in the treatment group, and 0.081 in the control group.

Indigenous people have 3.7% service business ownership as opposed to 15.3% in the reference population [@businessesper]. They are 0.24 times as likely to own a service business in Ontario.
Therefore, we can estimate that 0.009 of business owners in the treatment group are Indigenous, and 0.011 in the control group.

## Survey 1 and 2

We randomly sampled to assign each restaurant to either "yes", "no", or "nonanswer", for the owner being Indigenous, a visible minority, or a woman, using the proportions above (with 0.01 set aside for "nonresponse"), in Table \@ref(tab:minority1) and Table \@ref(tab:minority2).

```{r minority1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(woman = sample(c("yes","no","nonanswer"), 
                        size = num_total, 
                        prob = c(0.154, 0.836, 0.01), 
                        replace = TRUE))

set.seed(19893)
survey_1 <- bind_rows(
  survey_1 %>%
    filter(group == "treatment") %>%
    mutate(indigenous = sample(c("yes", "no", "nonanswer"), 
                               size = num_treat, 
                               prob = c(0.009, 0.981, 0.01), 
                               replace = TRUE))
  ,
  survey_1 %>%
    filter(group == "control") %>%
   mutate(indigenous = sample(c("yes", "no", "nonanswer"), 
                              size = num_control, 
                              prob = c(0.011, 0.979, 0.01), 
                              replace = TRUE))
)

set.seed(19893)
survey_1 <- bind_rows(
  survey_1 %>%
    filter(group == "treatment", indigenous != "yes") %>%
    mutate(visible = sample(c("yes", "no", "nonanswer"), 
                               size = first(count(survey_1 %>% 
                                                    filter(group == "treatment", 
                                                           indigenous != "yes"))), 
                               prob = c(0.057, 0.933, 0.01), 
                            replace = TRUE))
  ,
  survey_1 %>%
    filter(group == "control", indigenous != "yes") %>%
   mutate(visible = sample(c("yes", "no", "nonanswer"), 
                              size = first(count(survey_1 %>% 
                                                   filter(group == "control", 
                                                          indigenous != "yes"))), 
                              prob = c(0.08, 0.91, 0.01), 
                           replace = TRUE))
  ,
  survey_1 %>%
    filter(indigenous == "yes") %>%
    mutate(visible = "no")
)

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner minority status and gender to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r minority2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(woman = sample(c("yes","no","nonanswer"), 
                        size = num_total, 
                        prob = c(0.154, 0.836, 0.01), 
                        replace = TRUE))

set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(indigenous = sample(c("yes", "no", "nonanswer"), 
                               size = num_treat, 
                               prob = c(0.009, 0.981, 0.01), 
                               replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
   mutate(indigenous = sample(c("yes", "no", "nonanswer"), 
                              size = num_control, 
                              prob = c(0.011, 0.979, 0.01), 
                              replace = TRUE))
)

set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment", indigenous != "yes") %>%
    mutate(visible = sample(c("yes", "no", "nonanswer"), 
                               size = first(count(survey_2 %>% 
                                                    filter(group == "treatment", 
                                                           indigenous != "yes"))), 
                               prob = c(0.057, 0.933, 0.01), 
                            replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control", indigenous != "yes") %>%
   mutate(visible = sample(c("yes", "no", "nonanswer"), 
                              size = first(count(survey_2 %>% 
                                                   filter(group == "control", 
                                                          indigenous != "yes"))), 
                              prob = c(0.08, 0.91, 0.01), replace = TRUE))
  ,
  survey_1 %>%
    filter(indigenous == "yes") %>%
    mutate(visible = "no")
)

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner minority status and gender to Survey 2 data") %>%
  kableExtra::kable_styling()
```

# Shutdowns

## Research

43.2% of restaurants shut down temporarily due to the pandemic [@covid19impactonbusiness].

3% closed of restaurants in the US closed permanently between March and June [@restaurantclosures]. This period was about six times as long as our two week closure.

# Survey 2

We randomly sampled to assign each restaurant to either "none", "temporary", or "permanent" for shutdowns, with probabilities of 0.558, 0.432, and 0.01, in Table \@ref(tab:closures).

```{r closures, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign shutdown using the ratios above only for the treatment group:

set.seed(19893)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"),
                             size = num_treat,
                             prob = c(0.558, 0.432, 0.01),
                             replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"),
                             size = num_control,
                             prob = c(.99, 0.005, 0.005),
                             replace = TRUE))
) %>%
  arrange(ID)

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding closures to Survey 2 data") %>%
  kableExtra::kable_styling()
```

# Revenue

## Research into restaurant revenue

In 2019, there were 25,836 restaurants and eating-places that were employers [@businesseswithemployees] and 6,968 restaurants that were not employers [@businesseswithoutemployees] in Canada, for a total of 32,804 restaurants.

In the same year, full-service restaurants had \$13,456,600,000 in revenue, and limited-service eating places had \$14,082,700,000 in revenue [@foodservicessummary].

From this, we can calculate an estimated \$69,959 average revenue per month.

From an analysis of Kaggle restaurant data [@kagglerestaurantdata], we can see that restaurant revenue looks like an F-distribution.
The distribution rf(n, df1, df2, ncp) = rf(num_rest, 20, 20) gives an approximately correct shape. Using the F-distribution formula, we know the distribution has a mean of 10/9, so we would multiply by 62,963 to get the mean to 69,959.

## Survey 1

We randomly sampled from the above F-distribution to generate revenues for the restaurants, in Table 
\@ref(tab:revenue1).

```{r revenue1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues:

set.seed(19894)
survey_1 <- survey_1 %>%
  mutate(revenue = 62963 * rf(num_total,20,20))

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding revenue to Survey 1 data") %>%
  kableExtra::kable_styling()
```

## Research into changes in revenue due to closures

First, the revenues from the first survey will have a randomized factor (normal distribution with a mean of 1 and a standard distribution of 0.1) to add some random variance between months.

Second, places that are closed permanently will be assumed to experience -75% average revenue, and temporary shutdowns will be assumed to experience -50% revenue.

Revenues went to -35% during the first lockdown wave in March [@bythenumbers]. We will assume that these numbers are accurate for a dine-in/takeout restaurant that is experiencing a takeout-only closure. They are closed 14 out of 31 days, so we will assume average losses of -17%.

We will assume that takeout-only restaurants will not have losses on average, and dine-in only places will have -100% revenue on average for the days they are closed. With closures 14 out of 31 days in the month, dine-in only will have average losses of -45%.

All businesses had a median decrease of -15%, but visible minority-owned businesses had a median decrease of -25% [@aug20192020]. Therefore, an additional -12% will be applied to visible minority-owned restaurants.

There seems to be no significant difference in business losses for disabled owners or female business owners [@aug20192020].

## Survey 2

We started by randomly sampling from the same F-distribution multiplied by 62963, but also multiplied by random sampling from normal distribution centred at 1, with a standard deviation of 0.1, to add some variation from the previous data. For the treatment group data, we then applied a factor that reduced revenue based on level of shutdown (0.5 for temporary and 0.25 for permanent), another factor that reduced revenue based on service type (0.55 for dine-in and 0.83 for both), and another factor that reduced revenue based on minority status (0.88 for minority business owners). This can be seen in Table \@ref(tab:revenue2).

```{r revenue2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues, 
# and apply a random number around 1 to add variation:

set.seed(19894)
survey_2 <- survey_2 %>%
  mutate(revenue = 62963 * rf(num_total,20,20)) %>%
  mutate(factor = abs(rnorm(num_total, 1, .1))) %>%
  mutate(revenue = factor * revenue) %>%
  select(-factor)

# Reduce the treatment revenues to those that shutdown:

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(shutdown == "none") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(shutdown == "temporary") %>%
      mutate(revenue = 0.50 * revenue)
    ,
    survey_2 %>%
      filter(shutdown == "permanent") %>%
      mutate(revenue = 0.25 * revenue)
  ) %>%
  arrange(ID)

# reduce the treatment revenues for those with more dine-in and less take-out service:

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(service_type == "dinein" & group == "treatment") %>%
      mutate(revenue = 0.55 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "both" & group == "treatment") %>%
      mutate(revenue = 0.83 * revenue)
    ,
    survey_2 %>%
      filter(service_type == "takeout" & group == "treatment") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(group == "control")
  ) %>%
  arrange(ID)

# reduce the treatment revenues further for racial and ethnic minority owned businesses:

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter((visible == "yes" | indigenous == "yes") 
             & group == "treatment") %>%
      mutate(revenue = 0.88 * revenue)
    ,
    survey_2 %>%
      filter(visible == "no" & group == "treatment")
    ,
    survey_2 %>%
      filter(group == "control")
  )

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding revenue to Survey 2 data") %>%
  kableExtra::kable_styling()
```

# Number of employees

## Research

Restaurants averaged $82,000 in yearly sales (gross revenue) per full-time equivalent employee [@restfact]. Therefore, we can estimate 6,833 monthly revenue per full-time equivalent employee.

In January 2021, there were 2187300 full time sales and service employees in Canada and 1164000 part time sales and service employees [@hourswages]. Therefore, there are 1.9 full time employees per part time employee.
Part time employees in service and sales averaged 16.6 hours per week [@hourswages]. We will assume that 2 PTE = 1 FTE, or 1 PTE = 0.5 FTE.
Therefore, for every 1.9 + 1 employees, there are 1.9 + 1*0.5 FTE. For every 2.9 employees, there are 2.4 FTE.

$1.9 / (1.9 + 0.5) = 1.9 / 2.4 = 0.79$

$1 / (1.9 + 0.5) = 1 / 2.4 = 0.42$

## Survey 1

We divided revenue by 6833, multiplied by 0.79 and rounded to find the number of full time employees, and multiplied by 0.42 and rounded to find the number of part time employees, in Table \@ref(tab:employ1).

```{r employ1, echo = FALSE, message = FALSE, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(fte = revenue / 6833) %>%
  mutate(ft = round(fte * 0.79)) %>%
  mutate(pt = round(fte * 0.42)) %>%
  select(-fte)

survey_1 <- bind_rows(
  survey_1 %>% filter(ft < 1) %>%
    mutate(ft = 1),
  survey_1 %>% filter(ft >= 1)
)

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding employment numbers to Survey 1 data") %>%
  kableExtra::kable_styling()
```

## Survey 2

For the second survey, we will assume that treatment restaurants shifted towards part time random amounts-- that is, the number of full time employees per part time employee dropped from 1.9, a small drop that varied per restaurant.

We divided revenue by 6833. We used the absolute value of random sample of a normal distribution centred at zero, with a standard deviation of 0.1, as a factor to decrease the part-time ratio of 1.9.
We then used the new ratio to calculate the number of full-time and part-time employees. This can be seen in Table \@ref(tab:employ2).

```{r employ2, echo = FALSE, message = FALSE, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

set.seed(19893)
survey_2 <- survey_2 %>%
  mutate(fte = revenue / 6833) %>%
  mutate(ptincrease = 1.9 * (1 - abs(rnorm(num_total, 0, sd = 0.1)))) %>%
  mutate(ft = round(fte * (ptincrease / (ptincrease + 0.5)))) %>%
  mutate(pt = round(fte * (1 / (ptincrease + 0.5)))) %>%
  select(-fte, - ptincrease)

survey_2 <- bind_rows(
  survey_2 %>% filter(ft < 1) %>%
    mutate(ft = 1),
  survey_2 %>% filter(ft >= 1)
)

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding employment numbers to Survey 2 data") %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

# Non-response

## Research

A professional mail survey with follow up found a non-response rate of 61% [@nonresponse].
We will also assume that some people will mail in answers and fill in the form incompletely.

## Survey 1 and 2

We randomly sampled to assign each row as "yes", "partial", or "no", with probabilities 0.6, 0.01, and 0.39, for nonresponse. We then discarded all the "no" rows. For the partial rows, we randomly selected a column to delete the data for. This can be seen in Table \@ref(tab:response1) and Table \@ref(tab:response2).

```{r response1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

num_cols_1 = 11

set.seed(891)
survey_1 <- survey_1 %>%
  mutate(nonresponse = sample(c("yes","partial","no"), 
                              size = num_total, 
                              prob = c(0.6, 0.01, 0.39), 
                              replace = TRUE)) %>%
  filter(nonresponse != "yes")

set.seed(891)
survey_1 <- survey_1 %>%
  mutate(which_partial = sample(c(4:num_cols_1), 
                                size = first(count(survey_1)), 
                                replace = TRUE))

for (i in c(1:as.numeric(count(survey_1)))){
  if (survey_1[i, "nonresponse"] == "partial") {
    survey_1[i, as.numeric(survey_1[i,"which_partial"])] <- NA
  }
}

survey_1 <- survey_1 %>%
  select(-nonresponse, -which_partial)

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding nonresponse to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r response2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

num_cols_2 = 12

set.seed(198)
survey_2 <- survey_2 %>%
  mutate(nonresponse = sample(c("yes","partial","no"), 
                              size = num_total, 
                              prob = c(0.6, 0.01, 0.39), 
                              replace = TRUE)) %>%
  filter(nonresponse != "yes")

set.seed(198)
survey_2 <- survey_2 %>%
  mutate(which_partial = sample(c(4:num_cols_1), 
                                size = first(count(survey_2)), 
                                replace = TRUE))

for (i in c(1:as.numeric(count(survey_2)))){
  if (survey_2[i, "nonresponse"] == "partial") {
    survey_2[i, as.numeric(survey_2[i,"which_partial"])] <- NA
  }
}

survey_2 <- survey_2 %>%
  select(-nonresponse, -which_partial)

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding nonresponse to Survey 2 data") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

```{r, include = FALSE}
# Save the two survey results tables:

write_csv(survey_1, here("outputs/data/survey_1.csv"))

write_csv(survey_2, here("outputs/data/survey_2.csv"))
```

# Budget

We Calculated our budget based on Canada Post's business costs [@canadapost], Ontario's minimum wage [@minimumwage], and pricing for data management systems for surveys [@alchemer]. The budget can be seen in Table \@ref(tab:budget).

```{r budget, echo = FALSE, message = FALSE, warning = FALSE}
# research for typical budget numbers
restaurants = 2006
survey_total = 2
# https://www.canadapost.ca/cpc/en/business/postal-services/mailing/letter-discounts.page?
indiv_postage = 0.85 
# https://www.ontario.ca/document/your-guide-employment-standards-act-0/minimum-wage
minimum_wage = 14.25 
# https://www.alchemer.com/plans-pricing/small-teams/
management_service = 1300 
average_call_mins = 7
response_rate = .3

# calculate costs
phone_interviewer_cost = 
  round(restaurants * survey_total*response_rate, 2) * average_call_mins / 60 * minimum_wage

postage_cost = (restaurants * 2) * indiv_postage

# save the information in a dataframe
survey_budget <- data.frame(
  Item = c ("Data Management", "Phone Interviewer Wages", "Postage", "TOTAL"),
  Cost = c (management_service, 
            phone_interviewer_cost, 
            postage_cost, 
            sum(management_service, phone_interviewer_cost, postage_cost)),
  Description = c ("Online survey management account fees",
                   "1,203 calls (7 minutes on average) at $14.25/hr",
                   "4,012 mailers at $0.85/ea",
                   "")
)

survey_budget %>%
  knitr::kable(booktabs = TRUE, caption = "Total survey budget") %>%
  kableExtra::kable_styling(latex_options = "hold_position")

#save the survey for presentation in the main report.
write_csv(survey_budget, here("outputs/data/survey_budget.csv"))
```

\clearpage

\newpage

# References